Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-11-28 doc/manuscript/translocation.qmd
Finish CMR spec

TODO:

- State model description
- State transition matrix for natural recruits
- State transition matrix for introduced individuals
- Parameter model description
- Priors
- Specification of the joint distribution (possible diagram of the DAG)

>>>>

The state transition matrix $\Psi_{t, i}$ contains the probabilities of individual $i$ transitioning from state $u_{i, t}$ (rows) to $u_{i, t+1}$ (columns) between primary period $t$ and $t+1$. 
For unintroduced individuals which naturally recruit, this matrix is given by:

$$
\Psi_{t, i} =
\begin{blockarray}{cccc}
  \text{Not recruited} & \text{Alive} & \text{Dead} \\
\begin{block}{(ccc)c}
  1 - \lambda_t & \lambda_t & 0 & \text{Not recruited} \\
  0 & \phi_t & 1 - \phi_t & \text{Alive} \\
  0 & 1 & 1 & \text{Dead} \\
\end{block}
\end{blockarray},
$$
where $\lambda_t$ is the probability of recruiting in time $t$ and $\phi_t$ is the probability of survival in time $t$.

For introduced individuals which have deterministic recruitment (they recruit when introduced), the state transition matrix is given by:

$$
\Psi_{t, i} =
\begin{blockarray}{cccc}
  \text{Not recruited} & \text{Alive} & \text{Dead} \\
\begin{block}{(ccc)c}
  1 - I_{t, i} & I_{t, i} & 0 & \text{Not recruited} \\
  0 & \phi_t & 1 - \phi_t & \text{Alive} \\
  0 & 1 & 1 & \text{Dead} \\
\end{block}
\end{blockarray},
$$
where $I_{t, i}$ is a known indicator function for whether individual $i$ was introduced in primary period $t$.

We allow recruitment probabilities to vary in time via random effects, such that:

$$\lambda_t = \text{logit}^{-1}(\alpha^{(\lambda)} + \epsilon^{(\lambda)}_t),$$

where $\alpha^{(\lambda)}$ is an intercept parameter and $\epsilon^{(\lambda)}_t$ is an adjustment for time $t$.

Survival probabilities also vary in time, and as a function of known covariates:

$$\phi_t = \text{logit}^{-1}(X^{(\phi)}_t \beta^{(\phi)} + \epsilon^{(\phi)}_t),$$

where $X^{(\phi)}_t$ is a row vector of known covariates, $\beta^{(\phi)}$ is a column vector of unknown coefficients, and $\epsilon^{(\phi)}_t$ is an adjustment for time $t$.

To complete the specification of the Bayesian model, we need to specify priors for all unknown parameters. 
The recruitment parameter priors were specified as follows:

$$\alpha^{(\lambda)} \sim N(0,1),$$
$$\sigma^{(\lambda)} \sim N_+(0,1),$$
$$\epsilon^{(\lambda)}_t \sim N(0, \sigma^{(\lambda)}),$$
for periods $t=1, ..., T$. 
Here $N$ represents the normal distribution and $N_+$ the half normal distribution with positive support.

Survival parameter priors were specified similarly as:

$$\beta^{(\phi)} \sim N(0, 1),$$
$$\sigma^{(\phi)} \sim N_+(0, 1),$$
$$\epsilon^{(\phi)}_t \sim N(0, \sigma^{(\phi)}),$$
for time $t=1, ..., T$.

The detection model coefficient vector also received a standard normal prior $\beta^{(p)} \sim N(0, 1)$.

We computed the likelihood of each individual capture history using the forward algorithm, and we estimated the latent states using the forward-backward algorithm (@zucchini2009hidden, @joseph2018). 
All of the code to specify and fit the model in Stan is available in the open source mrmr package (@joseph2019).

The joint distribution of the resulting model can be written as follows:

$$
\begin{multline*}
[\alpha^{(\lambda)}, \sigma^{(\lambda)}, \epsilon^{(\lambda)}_{1:T}, \beta^{(\phi)}, \sigma^{(\phi)}, \epsilon^{(\phi)}_{1:T}, \beta^{(p)} \mid \pmb Y] \propto \prod_{i=1}^M [Y_i \mid \alpha^{(\lambda)}, \epsilon^{(\lambda)}_{1:T}, \beta^{(\phi)}, \epsilon^{(\phi)}_{1:T}, \beta^{(p)}] \times \\
\prod_{t=1}^T [\epsilon_t^{(\lambda)} \mid \sigma^{(\lambda)}] [\epsilon_t^{(\phi])} \mid \sigma^{(\phi)}] [\sigma^{(\lambda)}] [\sigma^{(\phi)}] [\alpha^{(\lambda)}] [\beta^{(\phi)}] [\beta^{(p)}],
\end{multline*}
$$

where $\pmb Y$ is an $M \times T$ detection matrix, and $Y_i$ the capture history of individual $i$.
